Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life

### 2021-01-06 100-phylota-overview.Rmd
intro clean up
intro

    16 families and 25 genera,and a tree of 33 genera from 20 different families

@@

    It does not cite phylota as it should, but includes it in their references.

>>>>

    16 families and 25 genera, and a tree of 33 genera from 20 different families

@@

    It does not reference to phylota in the main text (they should bc they used it), but includes it in their references.

### 2021-01-06 sources/bmc_template_submission/bmc_article.tex
intro

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Phylogenies capture the shared history of organisms and provide key evolutionary
context for our biological observations \cite{dobzhansky1973nothing}.
A challenge for phylogenetic reconstruction
is assembling high-quality homology hypotheses.
While genomics has, and will continue to, revolutionize phylogenetic inference,
the diversity of alternative genomic sequencing approaches implemented produce
largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
creating challenges for phylogenetic reconstruction.

While phylogenomics tackles this problem by focusing on targeted capture of informative
regions \cite{andermann2020guide}, it is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. It does not contribute to
democratization of research yet \cite{li2017enabling}.

Decades of single locus sequencing have generated
massive amounts of homologous DNA datasets that can be used for phylogenetic
reconstruction at many scales.
More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
million DNA sequences that were gauged to have the potential to resolve phylogenetic
relationships of 98.05\% of the almost 241,000
distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
The Open Tree of Life project (OpenTree)
constructs a comprehensive tree of life by synthesizing published phylogenies
and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
placed in the tree based on taxonomy. Meaning that there is a lot of information

Is genomic data on genbank available for a smaller percentage of all life, compared to
single loci in the same database?

Notable examples of phylogenies constructed using multi-locus, database mined, super matrix approach:
Magallon, Soltis, examples with fish, bats,

@@

Make emphasis on these really great studies (and pipelines) that do not rely on
genomic data, and that there is a lot of single locus data that can be analysed

@@

Public biological databases constitute an amazing resource for evolutionary
studies. Updating existing phylogenies with molecular data that has never been
incorporated into any phylogenetic estimate, geographical location, fossils, and
other data in a reproducible and continuous manner is possible by establishing a
data interoperability framework for biological databases. Here, we introduce
Physcraper, a tool that automates database connections to build upon homology
hypotheses that taxon specialists have assessed and deemed appropriate for a
specific phylogenetic scope to update a starting tree and single locus alignments
with public DNA data.

>>>>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Phylogenies capture the shared history of organisms and provide key evolutionary
context for our biological observations \cite{dobzhansky1973nothing}.
The Open Tree of Life project (OpenTree)
constructs a comprehensive tree of life by synthesizing published phylogenies
and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
placed in the tree based on taxonomy. Meaning that there is still a lot of work
to do to complete a full phylogenetic tree of all life.

Decades of single locus sequencing have generated
massive amounts of homologous DNA datasets that have the potential to be used for phylogenetic
reconstruction at many scales \cite{sanderson2008phylota}.
More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
million DNA sequences that were gauged to have the potential to resolve phylogenetic
relationships of 98.05\% of the almost 241,000
distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.

With more than XXXX million single locus
sequences in GenBank today, there is a considerable amount of phylogenetically
informative data in GenBank that has not been analysed and/or made publicly available
with the potential to give phylogenetic light to the tree of life.

Phylogenies constructed from single locus super mayrix datasets are still widely
generated and relevant to the field. Notable examples include the

Notable examples of phylogenies constructed using multi-locus, database mined, super matrix approach:
Magallon, Soltis, examples with fish, bats,
\cite{magallon2015metacalibrated} mining by hand - flowering plant families - has been cited 561 times
\cite{smith2018constructing} 38k clades of plants with seeds - cited 176 times
\cite{jetz2012global} 6k bird species - cited
\cite{rabosky2018inverse} mining with phlawd - cited 202 times

@@

Make emphasis on these really great studies (and pipelines) that do not rely on
genomic data, and that there is a lot of single locus data that can be analysed.

@@
